---
title: "**int3ract**: Johnson--Neyman Technique and Its Three-Way Extension for Frequentist and Bayesian Models in R"
output-dir: docs
format:
  html:
    toc: true
    output-file: index.html
    number-sections: true
    embed-resources: true
author:
  - name: "Robert W. Krause"
    email: "robert.w.krause@mailbox.org"
abstract: |
  Interaction effects are ubiquitous in applied statistical modelling, yet
  their meaningful interpretation remains challenging. The classic
  Johnson--Neyman (JN) technique [@Johnson1936] addresses this challenge for
  two-way interactions by identifying the regions of a moderator's range over
  which a focal effect is and is not statistically significant. The
  **int3ract** package for R implements the JN technique and its three-way
  extension (the Johnson--Neyman--Krause, or JNK, technique) for both
  frequentist and Bayesian models. The function `JNK_freq()`
  auto-detects models fitted via `lm()`/`glm()`, **RSiena**'s `siena()`,
  or **lme4**'s `lmer()`/`glmer()`, but can also be applied to
  multiplicative interactions from (virtually) any model family by
  supplying a coefficient vector and covariance matrix directly.
  For Bayesian Stochastic Actor-Oriented Models (SAOMs) estimated with
  **multiSiena**, or any model producing posterior draws, the function
  `JNK_bayes()` produces conditional posterior distributions.
  For two-way interactions, classic shaded
  confidence-band plots  are created that visually demarcate significant and
  non-significant regions along the moderator range; three-way interactions
  yield colour-gradient heatmaps with optional crosshatch overlays for
  non-significant regions. The package is designed to encourage richer,
  region-specific reporting of interaction effects in place of the
  conventional single-slope spotlight approach.
keywords: [interaction effects, Johnson-Neyman technique, moderation analysis,
           SAOM, Bayesian, R]
keywords-formatted: [interaction effects, Johnson--Neyman technique,
                     moderation analysis, SAOM, Bayesian, "[R]{.proglang}"]
bibliography: int3ract_paper.bib
header-includes:
  - \usepackage{float}
---


```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Introduction {#sec-intro}

Multiplicative interaction terms allow researchers to examine whether the
relationship between a predictor $X$ and an outcome $Y$ depends on the value
of one or more additional variables---the so-called moderators. Interpreting
such interactions by inspecting the sign and significance of the interaction
coefficient alone is, however, insufficient and potentially misleading
[@Bauer2005; @Preacher2006]. The key insight behind the Johnson--Neyman (JN)
technique is that instead of evaluating the focal effect at a small number of
selected moderator values (often referred to as "spotlight" or "simple-slope"),
one can derive the *entire* region of the moderator's distribution over which
the focal effect is statistically distinguishable from zero [@Johnson1936;
@Potthoff1964].

Originally proposed in the context of the analysis of covariance
[@Johnson1936], the JN technique was later adapted for regression
[@Potthoff1964] and updated by @Bauer2005, whose
formulation is closest to what **int3ract** implements. Several R packages
already address the two-way JN case, for instance,  **interactions** [@Long2019]
and **jtools** [@Long2020] but none does so for three-way interactions. The
contribution of **int3ract** is threefold.

1. **Three-way interactions.** The package provides a complete implementation
   of the JNK technique for three-way multiplicative interactions, producing a
   heatmap over the two-dimensional moderator grid rather than a single-axis 
   ribbon plot.

2. **Model-agnostic frequentist engine.** `JNK_freq()` auto-detects `lm`,
   `glm`, `sienaFit`, and `lmerMod`/`glmerMod` objects, but also accepts a
   raw coefficient vector and covariance sub-matrix. It therefore applies to
   any model that produces these quantities -- ordinary least
   squares, generalised linear models, mixed-effects models, survival models,
    etc. -- without modification. For **lme4**
   models with random interaction terms, per-group Johnson--Neyman plots are
   available via `fixed_only = FALSE`.

3. **Bayesian models.** The function `JNK_bayes()` adapts the technique to
   the posterior distribution rather than point estimates, producing
   density plots (two-way) and posterior-mean heatmaps with Bayesian
   $p$-value overlays (three-way). It accepts **multiSiena** objects directly
   or raw posterior-draw matrices from any Bayesian estimation.

The remainder of the paper is organised as follows. @sec-background reviews
the statistical background of the JN technique and its three-way extension.
@sec-package describes the package architecture. @sec-freq and @sec-bayes
illustrate the frequentist and Bayesian workflows with worked examples. @sec-SAOM introduces the application to Stochastic Actor-Oriented Models (SAOMs) estimated with **RSiena** and **multiSiena**. @sec-graphics discusses graphical customisation. @sec-conc concludes.

## Statistical Background {#sec-background}

### The Two-Way Johnson--Neyman Technique

Consider the linear model

$$
Y_i = \beta_0 + \beta_1 X_i + \beta_2 M_i + \beta_3 X_i M_i + \varepsilon_i,
$$ {#eq-2way}

where $X$ is the focal predictor, $M$ is the moderator, and $\beta_3$ is the
interaction coefficient. The simple slope of $Y$ with respect to $X$ at a
given value $m$ of the moderator is

$$
\hat{\theta}_X(m) = \hat{\beta}_1 + \hat{\beta}_3 \, m.
$$ {#eq-simpleslope}

Its sampling variance follows from the delta method:

$$
\widehat{Var}\bigl[\hat{\theta}_X(m)\bigr]
  = \widehat{Var}(\hat{\beta}_1)
    + 2m\,\widehat{Cov}(\hat{\beta}_1, \hat{\beta}_3)
    + m^2\,\widehat{Var}(\hat{\beta}_3).
$$ {#eq-var2way}

The two-way JN regions are the set of moderator values $m$ for which the
two-sided $z$-test

$$
z(m) = \frac{\hat{\theta}_X(m)}
            {\sqrt{\widehat{Var}[\hat{\theta}_X(m)]}}
$$

satisfies $|z(m)| > z_{\alpha/2}$, where $z_{\alpha/2}$ is the
$(\alpha/2)$-upper quantile of the standard normal distribution. 

The symmetry of @eq-2way implies that the roles of $X$ and $M$ are
interchangeable: the simple slope of $Y$ on $M$ moderated by $X$ is
$\hat{\beta}_2 + \hat{\beta}_3 \, x$, with an analogous variance formula.
**int3ract** therefore always produces *two* JN plots per two-way
interaction and three for three-way interactions -- one for each variable acting as the focal predictor.

### The Three-Way Extension {#sec-threeway}

When a third variable $W$ enters as an additional moderator, the full
three-way interaction model is

$$
\begin{aligned}
Y_i = \; & \beta_0 + \beta_1 X_i + \beta_2 M_i + \beta_3 W_i
           + \beta_4 X_i M_i + \beta_5 X_i W_i + \beta_6 M_i W_i \\
         & + \beta_7 X_i M_i W_i + \varepsilon_i.
\end{aligned}
$$ {#eq-3way}

The conditional effect of $X$ given $M$ and $W$ is

$$
\theta_X(m, w)
  = \beta_1 + \beta_4 m + \beta_5 w + \beta_7 mw,
$$ {#eq-theta3way}

and the corresponding standard error is obtained via the multivariate delta
method:

$$
\begin{aligned}
\widehat{Var}\bigl[\hat{\theta}_X(m,w)\bigr]
  = \; & \widehat{Var}_{\beta_1}
      + m^2 \widehat{Var}_{\beta_4}
      + w^2 \widehat{Var}_{\beta_5}
      + (m w)^2 \widehat{Var}_{\beta_7} \\
    & + 2m \widehat{Cov}(\beta_1,\beta_4)
      + 2w \widehat{Cov}(\beta_1,\beta_5)
      + 2m w \widehat{Cov}(\beta_1,\beta_7) \\
    & + 2m w \widehat{Cov}(\beta_4,\beta_5)
      + 2m^2 w \widehat{Cov}(\beta_4,\beta_7) \\
    & + 2m w^2 \widehat{Cov}(\beta_5,\beta_7).
\end{aligned}
$$ {#eq-se3way}


Because the significance region is now two-dimensional, it cannot be depicted
as a simple interval on a line. **int3ract** instead renders a heatmap of
conditional point estimates $\hat{\theta}_X(m, w)$ using a diverging
colour scale, and superimposes a crosshatch pattern over cells where the null
hypothesis $\theta_X = 0$ should be rejected given the chosen $\alpha$-threshold
(default: `threshold = 0.05`).

## Package Architecture {#sec-package}

The package is organised around two exported functions and a shared utility
layer, summarised in @tbl-hierarchy.

| Function | Accepted input | Internal engine |
|---|---|---|
| `JNK_freq()` | `lm`, `glm`, `sienaFit`, `lmerMod`/`glmerMod`, or raw `covar`/`coefs` | `.freq_core()`, `parFun()`, `seFun()` |
| `JNK_bayes()` | `multiSiena`/`sienaBayesFit`, or raw posterior-draw matrix | `jnb_support2()`, `jnb_support3()` |

: Function hierarchy of **int3ract**. Both exported functions auto-detect the
model class and extract the relevant quantities before delegating to internal
engines in `utils.R`. {#tbl-hierarchy}

**Utility layer (`utils.R`).**
All internal helpers reside in a single file. `parFun()` and `seFun()` implement
@eq-theta3way and @eq-se3way. Model-specific extractors (`.extract_lm()`,
`.extract_siena()`, `.extract_lme4()`) handle the class-specific logic for
`JNK_freq()`. The Bayesian support functions `jnb_support2()` and
`jnb_support3()` produce the posterior-based plots for `JNK_bayes()`.

**Frequentist engine (`JNK_freq()`).**
`JNK_freq()` auto-detects the class of its first argument and extracts
coefficients, a covariance sub-matrix, variable names, and moderator value
ranges. For `lm`/`glm` and `lmerMod`/`glmerMod` objects, interaction-term names
are resolved automatically regardless of colon ordering. 
Researchers who can extract a coefficient vector and covariance matrix from any
other model (whether fitted by `coxph()`, `betareg()`, or any other estimator)
can call `JNK_freq()` directly by supplying `covar` and `coefs`.

For **lme4** models with random interaction terms, setting `fixed_only = FALSE`
produces per-group Johnson--Neyman plots using the group-specific coefficients
(fixed effects + BLUPs), mirroring the group-level output of `JNK_bayes()`.

**Bayesian engine (`JNK_bayes()`).**
For Bayesian models there is no single-point estimate or analytic covariance
matrix. Instead, conditional posteriors are used. `JNK_bayes()` accepts either
a raw matrix of posterior draws or a `multiSiena`/`sienaBayesFit` object 
(handling burn-in removal, thinning, and detection of fixed vs. random 
parameters).

For two-way interactions, `JNK_bayes()` computes the conditional posterior
draws

$$
\tilde{\theta}_X^{(s)}(m) = \theta^{(s)}_X + \theta^{(s)}_{XM} \cdot m
$$

at each sampled iteration $s$ and moderator value $m$, yielding a full
conditional posterior distribution. These are visualised as overlapping density
curves coloured by moderator value, and summarised by the posterior mean,
posterior standard deviation, and Bayesian $p$-value
$\Pr(\tilde{\theta}_X > 0 \mid \text{data}, \text{model}, \text{prior})$.

For three-way interactions, this is extended to the grid
$(m, w)$, producing heatmaps of both the posterior mean and the Bayesian
$p$-value, each overlaid with a crosshatch on cells outside the user-specified
"significance" thresholds.

## Frequentist Usage {#sec-freq}

### Two-Way Interaction with `lm`

The simplest entry point is `JNK_freq()` applied to an ordinary least-squares
model.

```{r}
#| label: example-2way-lm
#| eval: true
#| cache: true
library(int3ract)

set.seed(1402)
dat <- data.frame(x = rnorm(100),
                  z = rnorm(100),
                  w = rnorm(100))

dat$y = dat$x + 0.5 * dat$z + -.5 * dat$w + 0.5 * dat$x *
  dat$z * dat$w + rnorm(100,sd = 4)

mod2 <- lm(y ~ x * z, data = dat)

jn2 <- JNK_freq(mod2,
                theta_1 = "x",
                theta_2 = "z",
                alpha   = 0.05)
```

```{r}
#| label: fig-jn2
#| eval: true
#| echo: false
#| cache: true
#| dependson: "example-2way-lm"
#| fig-cap: >
#|   Johnson--Neyman plots for the two-way interaction between `x` and `z`
#|   ($\alpha = 0.05$). The ribbon shows pointwise 95% confidence bands for
#|   the conditional effect; green shading indicates regions where the effect
#|   is statistically significant, chocolate where it is not.
#| fig-subcap:
#|   - "Conditional effect of `x` as a function of the moderator `z`."
#|   - "Conditional effect of `z` as a function of the moderator `x`."
#| layout-ncol: 2
#| fig-width: 4
#| fig-height: 3.5
jn2$plots$x
jn2$plots$z
```

The returned list `jn2` contains `$param_table`, a list of two data frames
(one per each variable involved in the interactions) with columns for the
moderator values (`mod_vals`), conditional parameter, its standard error, and
$p$-value (`theta_vals`, `theta_se`, and `theta_p`), and a logical whether it is
significant given the chosen `threshold`; and `$plots`, a list of two
**ggplot2** objects.

In each plot the ribbon is filled with `sig_color` (default `'seagreen3'`)
where $|z| > z_{\alpha/2}$ and with `non_sig_color` (default `'chocolate'`)
elsewhere.

### Three-Way Interaction with `lm`

```{r}
#| label: example-3way-lm
#| eval: true
#| cache: true
#| dependson: "example-2way-lm"
mod3 <- lm(y ~ x * z * w, data = dat)

jn3 <- JNK_freq(mod3,
                theta_1    = "x",
                theta_2    = "z",
                theta_3    = "w",
                range_size = 100)

```

```{r}
#| label: fig-jn3
#| eval: true
#| echo: false
#| cache: true
#| dependson: "example-3way-lm"
#| fig-cap: >
#|   Johnson--Neyman--Krause heatmaps for the three-way interaction among
#|   `x`, `z`, and `w` (`range_size = 50`, $\alpha = 0.05$). Each cell
#|   shows the conditional point estimate of the focal effect across the
#|   two-dimensional moderator grid; the diverging colour scale runs from
#|   blue (negative) through yellow (zero) to red (positive).
#|   Crosshatched cells indicate regions where the conditional effect is
#|   not statistically significant.
#| fig-subcap:
#|   - "Conditional effect of `x` across the (`z`, `w`) moderator grid."
#|   - "Conditional effect of `z` across the (`x`, `w`) moderator grid."
#|   - "Conditional effect of `w` across the (`x`, `z`) moderator grid."
#| layout: [[1,1],[1]]
#| fig-width: 4
#| fig-height: 3.5
jn3$plots$x
jn3$plots$z
jn3$plots$w
```

Setting `range_size = 50` produces a $50 \times 50$ grid for each of the
three heatmaps (i.e., $3 \times 2{,}500$ conditional effect evaluations).
When all three moderator ranges contain fewer than eight distinct values, the
point estimate is printed in each cell; otherwise only the colour gradient
communicates the magnitude. The ranges for each variable are by default
extracted from the `lm`/`glm`/`lmerMod` object but can also be supplied
directly, which is always necessary for `sienaFit` objects and the generic
`covar`/`coefs` path (see the code chunks below).

### Mixed-Effects Models (**lme4**)

`JNK_freq()` natively supports `lmerMod` and `glmerMod` objects from **lme4**.
For fixed effects the interface is identical to `lm`:

```{r}
#| label: example-lme4
#| eval: false
library(lme4)
library(int3ract)

lme_mod <- lmer(y ~ x * z + (x * z | group), data = dat)

jn_lme <- JNK_freq(lme_mod,
                   theta_1 = "x",
                   theta_2 = "z")
```

When the interaction terms are random, setting `fixed_only = FALSE` produces
per-group Johnson--Neyman plots using group-specific coefficients (fixed
effects + BLUPs). The grouping variable defaults to the first random-effects
term but can be specified via `group_var`:

```{r}
#| label: example-lme4-groups
#| eval: false
jn_groups <- JNK_freq(lme_mod,
                      theta_1    = "x",
                      theta_2    = "z",
                      fixed_only = FALSE,
                      group_var  = "group")
```

### Direct Use with Coefficient Vectors

Because `JNK_freq()` also accepts a raw coefficient vector and covariance
matrix, it is straightforward to apply it to models not covered by the
built-in extractors or to Bayesian models when the posterior is reduced to a
covariance matrix and point estimates:

```{r}
#| label: example-generic
#| eval: false
JNK_freq(covar = my_vcov[c("x","z","x:z"), c("x","z","x:z")],
         coefs = my_coefs[c("x","z","x:z")],
         name  = c("x", "z"),
         theta_1 = "x", theta_2 = "z",
         theta_1_vals = c(-3, 3),
         theta_2_vals = c(-3, 3))
```

The same pattern applies to survival models (**survival**), ordinal regression
(**MASS**), beta regression (**betareg**), and any other estimation that
provides parameter estimates and their covariances.


## Bayesian Usage {#sec-bayes}

Using `JNK_bayes()` is as straightforward as using `JNK_freq()`. The function either expects a posterior matrix or sienaBayesFit/multiSiena object, from which it extracts the relevant posterior draws. The same arguments for specifying the parameters and their values apply as for `JNK_freq()`, but instead of confidence intervals and $p$-values, the output includes conditional posterior means, standard deviations, and Bayesian $p$-values 
($\Pr(\tilde{\theta}_X > 0 \mid \text{data}, \text{model}, \text{prior})$).

Parameters can either be specified by name (here, the posterior matrix `x` should be named accordingly) or by integer index (the position of the parameter in the posterior matrix). If names are given, only `theta_1`, `theta_2`, and, if applicable, `theta_3` need to be specified and the interactions are detected automatically (interactions should be named with the standard $\textsf{R}$ notation for interactions, i.e., `x:z`). 

The `thresholds` argument (a vector of length 2, e.g., `c(0.05, 0.95)`)
determines which cells of the posterior-mean heatmap are crosshatched: cells
whose Bayesian $p$-value falls inside the interval
$[\mathrm{threshold}_1, \mathrm{threshold}_2]$ can be treated as
`non-significant'. 

The function also handled burn-in iterations and thinning if desired via the the `thin` and `burn_in` arguments (default no thinning and no burn-in iterations).

```{r}
#| label: example-2way-bayes
#| eval: true
#| cache: true
#| dependson: "example-2way-lm"

library(MCMCpack)

mod_bayes2 <- MCMCregress(y ~ x * z, data = dat,
                          burnin = 1000, mcmc = 10000, thin = 1,
                          verbose = 0)


jnk_bayes2 <- JNK_bayes(mod_bayes2, theta_1 = 'x', theta_2 = 'z',
                        theta_1_vals = seq(-3,3,0.5), 
                        theta_2_vals = seq(-3,3,0.5))
```

```{r}
#| label: fig-jn-bayes2
#| eval: true
#| echo: false
#| cache: true
#| dependson: "example-2way-bayes"
#| fig-cap: >
#|   Johnson--Neyman--Krause plots for the two-way interaction between `x` and
#|   `z`. Each curve represents a conditional posterior distribution of the
#|   focal effect at a given value of the moderator.
#| fig-subcap:
#|   - "Conditional effect of `x` as a function of the moderator `z`."
#|   - "Conditional effect of `z` as a function of the moderator `x`."
#| layout-ncol: 2
#| fig-width: 4
#| fig-height: 3.5
jnk_bayes2$x_plot
jnk_bayes2$z_plot
```

Three-way plots in turn return heatmaps of the posterior mean and Bayesian $p$-value.

```{r}
#| label: example-3way-bayes
#| eval: true
#| cache: true
#| dependson: "example-2way-lm"

mod_bayes3 <- MCMCregress(y ~ x * z * w, data = dat,
                          burnin = 1000, mcmc = 10000, thin = 1,
                          verbose = 0)


jnk_bayes3 <- JNK_bayes(mod_bayes3, 
                        theta_1 = 'x', 
                        theta_2 = 'z',
                        theta_3 = 'w',
                        theta_1_vals = seq(-3,3,0.1), 
                        theta_2_vals = seq(-3,3,0.1), 
                        theta_3_vals = seq(-3,3,0.1),
                        thresholds = c(0.05, 0.95))
```

```{r}
#| label: fig-jn-bayes3
#| eval: true
#| echo: false
#| cache: true
#| dependson: "example-3way-bayes"
#| fig-cap: >
#|   Johnson--Neyman--Krause plots for the three-way interaction between `x`, 
#|   `z`, and `w`. The color gradient of the heatmap shows the value of
#|    conditional posterior mean of the focal effect. Crosshatch overlays
#|     indicate cells where the Bayesian $p$-value falls outside the threshold
#|      (here $p$ < 0.05 or $p$ > 0.95).
#| fig-subcap:
#|   - "Conditional posterior means of `w` as a function of the
#|    moderators `x` and `z`."
#|   - "Conditional posterior means of `z` as a function of the
#|    moderator `x`."
#| layout-ncol: 2
#| fig-width: 4
#| fig-height: 3.5
jnk_bayes3$post_mean_plots$w
jnk_bayes3$bayes_p_plots$w
```



## Stochastic Actor-Oriented Models {#sec-SAOM}

Interactions in SAOMs, estimated by **RSiena**'s `siena()` (formerly 
`siena07()`), between network statistics and actor attributes are common yet 
difficult to interpret [@Snijders26]. Introducing SAOMs is out of the scope for 
this paper. If you wish to learn more about them, see for 
instance @Snijders2017. In short, SAOMs are models for network evolution, 
designed to identify the structural and covariate drivers that lead to the 
creation, dissolution, and maintenance of connections in networks.

The same `JNK_freq()` function handles `sienaFit`
objects, using integer indices referencing the position of each effect in the
**RSiena** effects object^[While each effect in **RSiena** is named, within a
given model, effects are not necessarily identifiable by their name but the same
name might appear multiple times for different dependent variables or effect
types, making secure unique indexing by integer much easier for users.]. In 
`JNK_bayes()` the rate functions are skipped in the indexing!
The moderator value ranges (`theta_1_vals`, `theta_2_vals`, `theta_3_vals`)
should reflect the empirical support of the corresponding change statistics.

`JNK_bayes()` automatically detects whether each required parameter has been
estimated as a fixed effect ($\eta$; shared across groups) or as a random effect
($\mu$; group-varying, with a hyper-parameter), and extracts the appropriate
posterior draws. When `hyper_only = FALSE` and at least one of the involved 
parameters is
randomly varying between groups, `JNK_bayes()` additionally loops over all
$G$ groups and produces group-specific plots and tables, returned under the key
`random_groups_effects`. This is useful for assessing heterogeneity of
interaction effects across networks. It is strongly recommended to use `save = 
TRUE` when running `hyper_only = FALSE` so that all plots are directly created 
and saved. While this extends run-time, it will make it much easier to inspect 
the figures.

## Graphical Output and Customisation {#sec-graphics}

All plots are standard **ggplot2** objects [@Wickham2016] and can be further
modified using **ggplot2**'s layer system. The crosshatch pattern for
non-significant regions is produced by **ggpattern** [@ggpattern2022].
@tbl-colors documents the colour and pattern parameters shared across all
plotting functions. The colours for high, mid, and low point for continuous
plots are taken from the `Zissou1` palette from the **wesanderson** package
[@wesanderson]. Direct use of palettes is not supported but users can always
overlay their own color palettes post-hoc by calling
`scale_fill_gradient2()`/`scale_color_gradient2()` (for more and alternatives
see `?ggplot2::scale_fill_gradient()`).

```{r}
#| label: tbl-colors
#| tbl-cap: "Colour and pattern parameters available in all **int3ract**
#|  plotting functions. Default colour values are highlighted in their
#|   respective colour."
#| echo: false
#| cache: true

library(gt)

# Hex background and contrasting foreground for each colour-valued row.
# Rows without a colour default (grid_density, grid_spacing,
# crosshatch_non_sig) get NA and are left unstyled.
bg_cols <- c("#43CD80", "#D2691E", "#000000",
             "#3B9AB2", "#EBCC2A", "#F21A00",
             "#000000",  NA,        NA,
             "#666666",  NA)

fg_cols <- c("#000000", "#FFFFFF", "#FFFFFF",
             "#000000", "#000000", "#FFFFFF",
             "#FFFFFF",  NA,        NA,
             "#FFFFFF",  NA)

params <- data.frame(
  Parameter = c("sig_color", "non_sig_color", "line_color",
                "color_low", "color_mid",     "color_high",
                "color_grid", "grid_density", "grid_spacing",
                "color_values", "crosshatch_non_sig"),
  Description = c(
    "Ribbon fill for significant regions (2-way)",
    "Ribbon fill for non-significant regions (2-way)",
    "Point-estimate line colour (2-way)",
    "Low end of diverging fill scale (heatmaps)",
    "Midpoint colour of diverging fill scale",
    "High end of diverging fill scale",
    "Crosshatch line colour",
    "Density of crosshatch lines",
    "Spacing between crosshatch lines",
    "Within-cell text colour (heatmaps, \u2264 7 values)",
    "If TRUE, crosshatch marks the non-significant region"
  ),
  Default = c("`seagreen3'", "`chocolate'", "`black'",
              "`#3B9AB2'",  "`#EBCC2A'",     "`#F21A00'",
              "`black'",    "0.01",         "0.1",
              "`grey40'",   "TRUE"),
  stringsAsFactors = FALSE
)

# Build the base gt table with monospace styling on Parameter and Default
tbl <- gt(params) |>
  tab_style(
    style    = cell_text(font = google_font("Source Code Pro"), size = "small"),
    locations = cells_body(columns = c(Parameter, Default))
  ) |>
  tab_style(
    style    = cell_text(font = google_font("Source Code Pro"), size = "small"),
    locations = cells_column_labels(columns = c(Parameter, Default))
  ) |>
  cols_width(
    Parameter   ~ px(160),
    Description ~ px(330),
    Default     ~ px(110)
  ) |>
  opt_row_striping()

# Apply per-row background + foreground to the Default cell
for (i in seq_len(nrow(params))) {
  if (!is.na(bg_cols[i])) {
    tbl <- tbl |>
      tab_style(
        style = list(
          cell_fill(color = bg_cols[i]),
          cell_text(color = fg_cols[i])
        ),
        locations = cells_body(columns = Default, rows = i)
      )
  }
}

# Add a thin border around the 'white' cell so it is visible
tbl <- tbl |>
  tab_style(
    style     = cell_borders(sides = "all", color = "#CCCCCC", weight = px(1)),
    locations = cells_body(columns = Default, rows = 5)   # row 5 = color_mid
  )

tbl
```
## Summary and Outlook {#sec-conc}

**int3ract** provides a unified interface for the Johnson--Neyman technique
and its three-way extension across a broad class of frequentist and Bayesian
models. The two exported functions---`JNK_freq()` for frequentist models and
`JNK_bayes()` for Bayesian models---auto-detect the input class and handle
extraction internally, while remaining fully usable with raw coefficient
vectors, covariance matrices, or posterior-draw matrices for models not
covered by the built-in extractors. Native support for **lme4** mixed-effects
models includes per-group Johnson--Neyman plots when interaction terms are
random. Future development plans include:

1. A consideration for FDR-correction.
2. Additional Bayesian model classes beyond **multiSiena**.

Researchers are encouraged to report not only the JN boundaries but also the
full plots as supplementary material, in keeping with calls for more
transparent and complete reporting of interaction effects^[To give due credit,
next to the already listed packages, **int3ract** also depends on **scales**,
**tidyr**, and **tibble** [@scales; @tidy; @tibble]. This paper was written 
using **gt**, **lme4**, **MCMCpack**, and of coruse **quarto** using **tinytex** [@gt; @lme4; @quarto]] [@Simonsohn2020].

## References {.unnumbered}

::: {#refs}
:::
