---
title: "**int3ract**: Johnson--Neyman Technique and Its Three-Way Extension for Frequentist and Bayesian Models in R"
output-dir: docs
format:
  pdf:
    keep-tex: true
    number-sections: true
    fig-pos: "H"
  html:
    toc: true
    output-file: index.html
    number-sections: true
    embed-resources: true
    fig-pos: "H"
author:
  - name: "Robert W. Krause"
    email: "robert.w.krause@mailbox.org"
abstract: |
  Interaction effects are ubiquitous in applied statistical modelling, yet
  their meaningful interpretation remains challenging. The classic
  Johnson--Neyman (JN) technique [@Johnson1936] addresses this challenge for
  two-way interactions by identifying the regions of a moderator's range over
  which a focal effect is and is not statistically significant. The
  **int3ract** package for R implements the JN technique and its three-way
  extension (the Johnson--Neyman--Krause, or JNK, technique) for both
  frequentist and Bayesian models. The core function `JNK_int()`
  can be applied to multiplicative interactions from (virtually) any model
  family. Dedicated
  wrappers (`JNK_lm()` and `JNK_siena()`) streamline the workflow for
  models fitted via `lm()`/`glm()` and **RSiena**'s `siena()`,
  respectively. For Bayesian Stochastic Actor-Oriented Models (SAOMs)
  estimated with **multiSiena**, the function `JNSienaBayes()` produces
  conditional posterior distributions. For two-way interactions, classic shaded
  confidence-band plots  are created that visually demarcate significant and
  non-significant regions along the moderator range; three-way interactions
  yield colour-gradient heatmaps with optional crosshatch overlays for
  non-significant regions. The package is designed to encourage richer,
  region-specific reporting of interaction effects in place of the
  conventional single-slope spotlight approach.
keywords: [interaction effects, Johnson-Neyman technique, moderation analysis,
           SAOM, Bayesian, R]
keywords-formatted: [interaction effects, Johnson--Neyman technique,
                     moderation analysis, SAOM, Bayesian, "[R]{.proglang}"]
bibliography: int3ract_paper.bib
header-includes:
  - \usepackage{float}
---


```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Introduction {#sec-intro}

Multiplicative interaction terms allow researchers to examine whether the
relationship between a predictor $X$ and an outcome $Y$ depends on the value
of one or more additional variables---the so-called moderators. Interpreting
such interactions by inspecting the sign and significance of the interaction
coefficient alone is, however, insufficient and potentially misleading
[@Bauer2005; @Preacher2006]. The key insight behind the Johnson--Neyman (JN)
technique is that instead of evaluating the focal effect at a small number of
selected moderator values (often referred to as "spotlight" or "simple-slope"),
one can derive the *entire* region of the moderator's distribution over which
the focal effect is statistically distinguishable from zero [@Johnson1936;
@Potthoff1964].

Originally proposed in the context of the analysis of covariance
[@Johnson1936], the JN technique was later adapted for regression
[@Potthoff1964] and updated by @Bauer2005, whose
formulation is closest to what **int3ract** implements. Several R packages
already address the two-way JN case, for instance,  **interactions** [@Long2019]
and **jtools** [@Long2020]. The contribution of **int3ract** is threefold.

1. **Three-way interactions.** The package provides a complete implementation
   of the JNK technique for three-way multiplicative interactions, producing a
   heatmap over the two-dimensional moderator grid rather than a single-axis 
   ribbon plot.

2. **Model-agnostic frequentist engine.** The internal function `JNK_int()`
   requires only a coefficient vector and the relevant sub-matrix of the
   parameter covariance matrix. It therefore applies to any frequentist
   model that produces these quantities---ordinary least squares, generalised
   linear models, mixed-effects models, survival models, or Stochastic
   Actor-Oriented Models---without modification.

3. **Bayesian SAOMs.** The function `JNSienaBayes()` adapts the technique to
   the posterior distribution rather than the sampling distribution, producing
   density plots (two-way) and posterior-mean heatmaps with Bayesian
   $p$-value overlays (three-way) for models estimated with **multiSiena**.

The remainder of the paper is organised as follows. @sec-background reviews
the statistical background of the JN technique and its three-way extension.
@sec-package describes the package architecture. @sec-freq and @sec-bayes
illustrate the frequentist and Bayesian workflows with worked examples.
@sec-graphics discusses graphical customisation. @sec-conc concludes.

## Statistical Background {#sec-background}

### The Two-Way Johnson--Neyman Technique

Consider the linear model

$$
Y_i = \beta_0 + \beta_1 X_i + \beta_2 M_i + \beta_3 X_i M_i + \varepsilon_i,
$$ {#eq-2way}

where $X$ is the focal predictor, $M$ is the moderator, and $\beta_3$ is the
interaction coefficient. The simple slope of $Y$ with respect to $X$ at a
given value $m$ of the moderator is

$$
\hat{\theta}_X(m) = \hat{\beta}_1 + \hat{\beta}_3 \, m.
$$ {#eq-simpleslope}

Its sampling variance follows from the delta method:

$$
\widehat{Var}\bigl[\hat{\theta}_X(m)\bigr]
  = \widehat{Var}(\hat{\beta}_1)
    + 2m\,\widehat{Cov}(\hat{\beta}_1, \hat{\beta}_3)
    + m^2\,\widehat{Var}(\hat{\beta}_3).
$$ {#eq-var2way}

The two-way JN regions are the set of moderator values $m$ for which the
two-sided $z$-test

$$
z(m) = \frac{\hat{\theta}_X(m)}
            {\sqrt{\widehat{Var}[\hat{\theta}_X(m)]}}
$$

satisfies $|z(m)| > z_{\alpha/2}$, where $z_{\alpha/2}$ is the
$(\alpha/2)$-upper quantile of the standard normal distribution. 

The symmetry of @eq-2way implies that the roles of $X$ and $M$ are
interchangeable: the simple slope of $Y$ on $M$ moderated by $X$ is
$\hat{\beta}_2 + \hat{\beta}_3 \, x$, with an analogous variance formula.
**int3ract** therefore always produces *two* JN plots per two-way
interaction and three for three-way interactions -- one for each variable acting as the focal predictor.

### The Three-Way Extension {#sec-threeway}

When a third variable $W$ enters as an additional moderator, the full
three-way interaction model is

$$
\begin{aligned}
Y_i = \; & \beta_0 + \beta_1 X_i + \beta_2 M_i + \beta_3 W_i
           + \beta_4 X_i M_i + \beta_5 X_i W_i + \beta_6 M_i W_i \\
         & + \beta_7 X_i M_i W_i + \varepsilon_i.
\end{aligned}
$$ {#eq-3way}

The conditional effect of $X$ given $M$ and $W$ is

$$
\theta_X(m, w)
  = \beta_1 + \beta_4 m + \beta_5 w + \beta_7 mw,
$$ {#eq-theta3way}

and the corresponding standard error is obtained via the multivariate delta
method:

$$
\begin{aligned}
\widehat{Var}\bigl[\hat{\theta}_X(m,w)\bigr]
  = \; & \widehat{Var}_{\beta_1}
      + m^2 \widehat{Var}_{\beta_4}
      + w^2 \widehat{Var}_{\beta_5}
      + (m w)^2 \widehat{Var}_{\beta_7} \\
    & + 2m \widehat{Cov}(\beta_1,\beta_4)
      + 2w \widehat{Cov}(\beta_1,\beta_5)
      + 2m w \widehat{Cov}(\beta_1,\beta_7) \\
    & + 2m w \widehat{Cov}(\beta_4,\beta_5)
      + 2m^2 w \widehat{Cov}(\beta_4,\beta_7) \\
    & + 2m w^2 \widehat{Cov}(\beta_5,\beta_7).
\end{aligned}
$$ {#eq-se3way}


Because the significance region is now two-dimensional, it cannot be depicted
as a simple interval on a line. **int3ract** instead renders a heatmap of
conditional point estimates $\hat{\theta}_X(m, w)$ using a diverging
colour scale, and superimposes a crosshatch pattern over cells where the null
hypothesis $\theta_X = 0$ should be rejected given the chosen $\alpha$-threshold
(default: `threshold = 0.05`).

## Package Architecture {#sec-package}

The package is organised into four layers, summarised in @tbl-hierarchy.

| Entry point | Model class | Core engine |
|---|---|---|
| `JNK_lm()` | `lm`, `glm` | `JNK_int()` |
| `JNK_siena()` | `sienaFit` (frequentist) | `JNK_int()` |
| `JNSienaBayes()` | `multiSiena` (Bayesian) | `jnb_support2()`, `jnb_support3()` |
| `JNK_int()` | *any* (generic) | `parFun()`, `seFun()` |

: Function hierarchy of **int3ract**. User-facing wrappers extract the relevant
coefficients and covariance sub-matrix from the model object and pass them to
the appropriate engine. {#tbl-hierarchy}

**Utility layer (`utils.R`).**  
`parFun()` and `seFun()` implement @eq-theta3way and @eq-se3way. 

**Core frequentist engine (`JNK_int()`).**  
The function `JNK_int()` accepts `name` (character vector of variable names),
`coefs` (named numeric vector of model coefficients), `covar` (the
corresponding covariance sub-matrix), and `vals` (a list of moderator value
ranges). When a 2-way interaction is analysed, it returns two ribbon plots and
two data frames; when a 3-way interaction, it returns three heatmaps together
with the underlying matrices of point estimates, standard errors, $p$-values,
and significance indicators.

Crucially, `JNK_int()` is model-agnostic. Any researcher who can extract a
coefficient vector and covariance matrix from their model (whether fitted by
`lm()`, `glmer()`, `coxph()`, or any other estimator) can call `JNK_int()`
directly. The function can also be applied to Bayesian estimations, if one 
would reduce the posteriors to their mean and covariance structure.

**Frequentist wrappers.**  
`JNK_lm()` extracts `coefficients` and `vcov()` from an `lm` or `glm` object
and constructs the interaction-term names following $\textsf{R}$'s standard `:` 
convention before delegating to `JNK_int()`.

`JNK_siena()` extracts `$theta` and `$covtheta` from a `sienaFit` object.
Because **RSiena** effects are harder to index by name^[While effects are 
identifyable by name, in complex models the same effect can appear multiple 
times for different dependent variables and/or as different effect types, making
 it harder on the user to provide accurate input to the `JNK_siena()` compared
  to using the already provided numbering system in `siena()` outputs.], the 
  user supplies the indices for all main effects and interaction terms. 

**Bayesian engine (`JNSienaBayes()`, `jnb_support2()`, `jnb_support3()`).**  
For Bayesian SAOMs there is no single-point estimate or analytic covariance
matrix. Instead, conditional posteriors will be used. `JNSienaBayes()` handles
burn-in removal and thinning, detects whether each parameter was estimated as
fixed ($\eta$) or random ($\mu$), and routes to `jnb_support2()` for 2-way
interactions or `jnb_support3()` for 3-way interactions.

For two-way interactions, `jnb_support2()` computes the conditional posterior
draws

$$
\tilde{\theta}_X^{(s)}(m) = \theta^{(s)}_X + \theta^{(s)}_{XM} \cdot m
$$

at each sampled iteration $s$ and moderator value $m$, yielding a full
conditional posterior distribution. These are visualised as overlapping density
curves coloured by moderator value, and summarised by the posterior mean,
posterior standard deviation, and Bayesian $p$-value
$\Pr(\tilde{\theta}_X > 0 \mid \text{data}, \text{model}, \text{prior})$.

For three-way interactions, `jnb_support3()` extends this to the grid
$(m, w)$, producing heatmaps of both the posterior mean and the Bayesian
$p$-value, each overlaid with a crosshatch on cells outside the user-specified
"significance" thresholds.

## Frequentist Usage {#sec-freq}

### Two-Way Interaction with `lm`

The simplest entry point is `JNK_lm()` applied to an ordinary least-squares
model.

```{r}
#| label: example-2way-lm
#| eval: true
#| cache: true
library(int3ract)

set.seed(1402)
dat <- data.frame(x = rnorm(100),
                  z = rnorm(100),
                  w = rnorm(100))

dat$y = dat$x + 0.5 * dat$z + -.5 * dat$w + 0.5 * dat$x * 
  dat$z * dat$w + rnorm(100,sd = 4)

mod2 <- lm(y ~ x * z, data = dat)

jn2 <- JNK_lm(mod2,
              theta_1 = "x",
              theta_2 = "z",
              alpha   = 0.05)
```

```{r}
#| label: fig-jn2
#| eval: true
#| echo: false
#| cache: true
#| dependson: "example-2way-lm"
#| fig-cap: >
#|   Johnson--Neyman plots for the two-way interaction between `x` and `z`
#|   ($\alpha = 0.05$). The ribbon shows pointwise 95% confidence bands for
#|   the conditional effect; green shading indicates regions where the effect
#|   is statistically significant, chocolate where it is not.
#| fig-subcap:
#|   - "Conditional effect of `x` as a function of the moderator `z`."
#|   - "Conditional effect of `z` as a function of the moderator `x`."
#| layout-ncol: 2
#| fig-width: 4
#| fig-height: 3.5
jn2$plots$x
jn2$plots$z
```

The returned list `jn2` contains `$param_table`, a list of two data frames
(one per each variable involved in the interactions) with columns for the
moderator values (`mod_vals`), conditional parameter, its standard error, and
$p$-value (`theta_vals`, `theta_se`, and `theta_p`), and a logical whether it is
significant given the chosen `threshold`; and `$plots`, a list of two
**ggplot2** objects.

In each plot the ribbon is filled with `sig_color` (default `'seagreen3'`)
where $|z| > z_{\alpha/2}$ and with `non_sig_color` (default `'chocolate'`)
elsewhere.

### Three-Way Interaction with `lm`

```{r}
#| label: example-3way-lm
#| eval: true
#| cache: true
#| dependson: "example-2way-lm"
mod3 <- lm(y ~ x * z * w, data = dat)

jn3 <- JNK_lm(mod3,
               theta_1    = "x",
               theta_2    = "z",
               theta_3    = "w",
               range_size = 50)
```

```{r}
#| label: fig-jn3
#| eval: true
#| echo: false
#| cache: true
#| dependson: "example-3way-lm"
#| fig-cap: >
#|   Johnson--Neyman--Krause heatmaps for the three-way interaction among
#|   `x`, `z`, and `w` (`range_size = 50`, $\alpha = 0.05$). Each cell
#|   shows the conditional point estimate of the focal effect across the
#|   two-dimensional moderator grid; the diverging colour scale runs from
#|   blue (negative) through yellow (zero) to red (positive).
#|   Crosshatched cells indicate regions where the conditional effect is
#|   not statistically significant.
#| fig-subcap:
#|   - "Conditional effect of `x` across the (`z`, `w`) moderator grid."
#|   - "Conditional effect of `z` across the (`x`, `w`) moderator grid."
#|   - "Conditional effect of `w` across the (`x`, `z`) moderator grid."
#| layout: [[1,1],[1]]
#| fig-width: 4
#| fig-height: 3.5
jn3$plots$x
jn3$plots$z
jn3$plots$w
```

Setting `range_size = 50` produces a $50 \times 50$ grid for each of the
three heatmaps (i.e., $3 \times 2{,}500$ conditional effect evaluations).
When all three moderator ranges contain fewer than eight distinct values, the
point estimate is printed in each cell; otherwise only the colour gradient
communicates the magnitude. The ranges for each variable are by default
extracted from the `lm`/`glm` object but can also be supplied directly, which is
always necessary for `JNK_int()` and `JNK_siena()` (see the code chunk below).

### Stochastic Actor-Oriented Models (`JNK_siena()`)

SAOMs estimated by **RSiena**'s `siena()` (formerly `siena07()`) are the primary
 application domain for which this package was developed, because interaction 
effects between network statistics and actor attributes are common yet difficult
to interpret [@Snijders26]. The interface mirrors `JNK_lm()` but uses integer 
indices referencing the position of each effect in the **RSiena** effects 
object^[While each effect in **RSiena** is named, within a given model, effects are not necessarily identifiable by their name but the same name might appear multiple times for different dependent variables or effect types].

```{r}
#| label: example-siena
#| eval: false
# Two-way interaction
x2 <- JNK_siena(res,
                 theta_1      = 6,
                 theta_2      = 7,
                 theta_int_12 = 8,
                 theta_1vals  = -10:10,
                 theta_2vals  = -10:10)

# Three-way extension
x3 <- JNK_siena(res,
                 theta_1       = 6,
                 theta_2       = 7,
                 theta_3       = 5,
                 theta_int_12  = 8,
                 theta_int_13  = 10,
                 theta_int_23  = 9,
                 theta_int_123 = 11,
                 theta_1vals   = -10:10,
                 theta_2vals   = -10:10,
                 theta_3vals   = 0:6,
                 range_size    = 10)
```

The moderator value ranges (`theta_1vals`, `theta_2vals`, `theta_3vals`)
should reflect the empirical support of the corresponding change statistics.
By default `range_size = 500` equidistant points are drawn from these ranges.
The returned tables and plots are named as derived from the model using the
`name' column in the **RSiena** effects object.

### Direct Use of `JNK_int()`

Because `JNK_int()` requires only a coefficient vector and a covariance
matrix, it is straightforward to apply it to models not covered by the
existing wrappers. For instance, for a mixed-effects model fitted with
**lme4**:

```{r}
#| label: example-lme4
#| eval: false
library(lme4)
library(int3ract)

lme_mod <- lmer(y ~ x * z + (1 | group), data = dat)

all_coef <- fixef(lme_mod)
all_vcov <- as.matrix(vcov(lme_mod))
sel      <- c("x", "z", "x:z")

JNK_int(name  = c("x", "z"),
         coefs = all_coef[sel],
         covar = all_vcov[sel, sel],
         vals  = list(range(dat$x), range(dat$z)),
         range_size = 1000)
```

The same pattern applies to survival models (**survival**), ordinal regression
(**MASS**), beta regression (**betareg**), and any other estimation that 
that provides parameter estimates and their covariances.

## Bayesian Usage: `JNSienaBayes()` {#sec-bayes}

For multi-group Bayesian SAOMs estimated with **multiSiena** 
[@SienaMult;@Kosk23], a series of conditional posterior distributions over the
range of the moderator replaces the classic JN-framework. If the moderator range
is given explicitly with fewer than eight unique values, distinct colors are
used to plot each The function
`JNSienaBayes()` automatically detects whether each required parameter has been
estimated as a fixed effect ($\eta$; shared across groups) or as a random effect
($\mu$; group-varying, with a hyper-parameter), and extracts the appropriate
posterior draws.

The returned object is a named list similar to those returned for frequentist 
interactions. Each table here contains the posterior mean, posterior SD,
Bayesian $p$-value
($\Pr(\tilde{\theta}_X > 0 \mid \text{data}, \text{model}, \text{prior})$), 
and 2.5th and 97.5th percentiles of the conditional posterior for each moderator
value.

When `hyper_only = FALSE` and at least one of the involved parameters is
randomly varying between groups, `JNSienaBayes()` additionally loops over all
$G$ groups and produces group-specific plots and tables, returned under the key
`random_groups_effects`. This is useful for assessing heterogeneity of
interaction effects across networks. Two additional arguments are helpful 
when many plots are made: `save = TRUE` will save each plot with `ggsave()`. 
To avoid cluttering your working directory, `folder` provides the location of 
a folder where the plots will be stored. If no input is provided, a new folder 
`int3ract JNplots` will be created (note that RStudio will ask you in the 
console whether such a folder is allowed to be created, if it does not exist 
yet).

For three-way interactions, the call is extended as usual. The
`thresholds` argument (a vector of length 2, e.g., `c(0.05, 0.95)`)
determines which cells of the posterior-mean heatmap are crosshatched: cells
whose Bayesian $p$-value falls inside the interval
$[\mathrm{threshold}_1, \mathrm{threshold}_2]$ are treated as
non-significant. 

```{r}
#| label: example-bayes-2way
#| eval: false
jnb2 <- JNSienaBayes(sbo          = my_multiSiena_result,
                     theta_1      = 4,
                     theta_2      = 7,
                     theta_int_12 = 9,
                     theta_1_vals = c(-1,0,1),
                     theta_2_vals = seq(-3,3,0.5),
                     burnIn       = 500,
                     thin         = 2,
                     save         = FALSE)
```


```{r}
#| label: fig-bayes-2fake
#| echo: false
#| eval: true
#| cache: true
#| fig-cap: >
#|   Conditional posterior densities for the parameters for reciprocity and
#|    altX on a attribute.
#| fig-subcap:
#|   - "Posterior densities for altX conditional on simultaneous changes in
#|    reciprocity. In this example, altX is always positive but has larger
#|    variance when reciprocity would decrease (i.e., a reciprocal tie to
#|    someone with negative values on the covariate would be broken,
#|    leading to -1 in recirpocity but overall an increase in altX)."
#|   - "Posterior densities for reciprocity conditional on simultaneous changes
#|    in altX. Here, while generally positive, reciprocity only has a
#|    consistently positive effect on tie formation when altX changes
#|    moderately, around -1 to 1 (yellow curves). While reciprocity gets more
#|    positive with increasing values in the change of altX, the posterior
#|    variance increases as well."
#| layout: [[1]]
#| fig-width: 4.5
#| fig-height: 3
set.seed(1402)
dat <- data.frame(x = rnorm(100),
                  z = rnorm(100),
                  w = rnorm(100))
dat$z <- ifelse(dat$z > 0,1,0)
dat$y = dat$x + 0.5 * dat$z + -1 * dat$w 

#+ 0.5 * dat$x * 
#  dat$z * dat$w + rnorm(100,sd = 4)

mod2 <- lm(y ~ x * z, data = dat)

jn2 <- JNK_lm(mod2,
              theta_1 = "x",
              theta_2 = "z",
              alpha   = 0.05)

theta <- mvtnorm::rmvnorm(50000, coefficients(mod2)[2:4],vcov(mod2)[2:4,2:4])
  
  

theta_1x <- as.vector(outer(theta[,3] , c(-1,0,1), 
                            FUN = '*')) + theta[,1]
theta_2x <- as.vector(outer(theta[,3] , seq(-3,3,0.5), FUN = '*')) + theta[,2]

t1plotData <- data.frame(modValue = as.factor(rep(round(c(-1,0,1),3),
                                                  each = nrow(theta))),
                         parameter = theta_1x)

t2plotData <- data.frame(modValue = as.factor(rep(round(seq(-3,3,0.5),3),
                                                  each = nrow(theta))),
                         parameter = theta_2x)
t1plotData$modValue <- as.numeric(as.character(t1plotData$modValue))
library(ggplot2)
g1 <- ggplot(t1plotData, aes(x = parameter,
                             group = modValue,
                       fill  = modValue,
                       color = modValue)) +
  geom_vline(xintercept = 0) +
  geom_density(alpha = 0.1, aes(fill = modValue, color = modValue)) +
  theme_bw() +
  xlab('altX') +
  ggplot2::scale_fill_gradient2(low  = "#3B9AB2",  mid = "#EBCC2A",
                                        high = "#F21A00", midpoint = 0) +
ggplot2::scale_color_gradient2(low  = "#3B9AB2",  mid = "#EBCC2A",
                                         high = "#F21A00", midpoint = 0) +
  ggtitle('Posterior distribution for altX \nconditional on Reciprocity') +  
  labs(color = 'Reciprocity Change',
       fill  = 'Reciprocity Change') 


g2 <- ggplot(t2plotData, aes(x = parameter,
                              group  = modValue,
                              fill   = as.numeric(as.character(modValue)),
                              color  = as.numeric(as.character(modValue)))) +
  geom_vline(xintercept = 0) +
  geom_density(alpha = 0.1) +
  ggplot2::scale_fill_gradient2(low  = "#3B9AB2",  mid = "#EBCC2A",
                                        high = "#F21A00", midpoint = 0) +
ggplot2::scale_color_gradient2(low  = "#3B9AB2",  mid = "#EBCC2A",
                                         high = "#F21A00", midpoint = 0) +
  theme_bw() +
  xlab('Reciprocity') +
  ggtitle('Posterior distribution for Reciprocity \nconditional on altX') +
  labs(color = 'altX \nChange \nStatistics',
       fill  = 'altX \nChange \nStatistics')
g1
g2
```




```{r}
#| label: example-bayes-3way
#| eval: false
jnb3 <- JNSienaBayes(sbo           = my_multiSiena_result,
                     theta_1       = 4,
                     theta_2       = 10,
                     theta_3       = 9,
                     theta_int_12  = 21,
                     theta_int_13  = 22,
                     theta_int_23  = 11,
                     theta_int_123 = 23,
                     theta_1_vals  = seq(-2, 2, length.out = 50),
                     theta_2_vals  = seq(-2, 2, length.out = 50),
                     theta_3_vals  = seq(-2, 2, length.out = 50),
                     thresholds    = c(0.05, 0.95),
                     burnIn        = 500,
                     thin          = 2)
```

```{r}
#| label: fig-bayes-3fake
#| echo: false
#| eval: true
#| cache: true
#| fig-cap: >
#|   Figures for Bayesian 3-way interactions
#| fig-subcap:
#|   "Mean posterior densities for the effect of ego's self-monitoring on tie
#|   creation conditional on ego's and alter's creativity. In this example,
#|   self-monitoring has a positive effect on tie creation when both ego and
#|   alter have similar levels of creativity."
#| layout: [[1]]
#| fig-width: 4.5
#| fig-height: 3
set.seed(1402)
library(ggplot2)
library(ggpattern)

dat <- data.frame(x = rnorm(100),
                  z = rnorm(100),
                  w = rnorm(100))
dat$y = dat$x + 0.5 * dat$z + -1 * dat$w + 0.5 * dat$x * 
  dat$z * dat$w + rnorm(100,sd = 4)

mod33 <- lm(y ~ x * z * w, data = dat)

set.seed(1402)
theta <- mvtnorm::rmvnorm(1000, 
                          coefficients(mod33)[2:8],
                          vcov(mod33)[2:8,2:8])

theta_1_vals <- seq(-2,2,length.out = 50)
theta_2_vals <- seq(-2,2,length.out = 50)
theta_3_vals <- seq(-2,2,length.out = 50)

theta_1n <- 'egoX-monitoring'
theta_2n <- 'altX-creativity'
theta_3n <- 'egoX-creativity'

thresholds <- c(0.05, 0.95)

nt <- nrow(theta)
n1 <- nt * length(theta_2_vals) * length(theta_3_vals)

grid <- expand.grid(mod1Val = theta_2_vals, mod2Val = theta_3_vals)

thetaX_matrix <- outer(rep(1, nt), rep(1, nrow(grid))) 
thetaX_matrix <- sapply(1:nrow(grid), function(k) {
  theta[,1] + grid$mod1Val[k] * theta[,4] +
    grid$mod2Val[k] * theta[,5] +
    grid$mod1Val[k] * grid$mod2Val[k] * theta[,7]
})

t1d <- data.frame(
  mod1      = rep(theta_2n, nrow(grid)),
  mod1Val   = rep(grid$mod1Val, each = nt),
  mod2      = rep(theta_3n, nrow(grid)),
  mod2Val   = rep(grid$mod2Val, each = nt),
  parameter = as.vector(thetaX_matrix)
)  |>
  dplyr::summarize(thetaPostMean = mean(parameter),
                   thetaPostSD = sd(parameter),
                   bayes_p = sum(parameter > 0) / length(parameter),
                   thetaPost2.5    = quantile(parameter,0.025),
                   thetaPost97.5   = quantile(parameter,0.975),
                   .by = c(mod1Val,mod2Val)
  )


t1d <- cbind(data.frame(theta = rep(theta_1n,nrow(t1d)),
                        mod1  = rep(theta_2n,nrow(t1d)),
                        mod2  = rep(theta_3n,nrow(t1d))),
             t1d)

t1d$sig <- t1d$bayes_p <= min(thresholds) |
  t1d$bayes_p >= max(thresholds)

t1d$pattern <- ifelse(!t1d$sig, "crosshatch",'none')

ns <- c(theta_1n,theta_2n,theta_3n)

color_mid = '#EBCC2A'
color_low = '#3B9AB2'
color_high = '#F21A00'
color_values = 'grey40'
color_grid = 'black'
grid_density = 0.01
grid_spacing = 0.1

pat <- t1d[!t1d$sig,]

plotsMean <- ggplot(t1d, aes(mod1Val, mod2Val)) +
  geom_tile(aes(fill = thetaPostMean )) +
  scale_color_identity() +
  scale_fill_gradient2(low = color_low, 
                       high = color_high,
                       mid = color_mid,
                       midpoint = 0) +
  geom_tile_pattern(data = pat, aes(pattern = pattern),
                    pattern_density = grid_density,
                    pattern_spacing = grid_spacing,
                    pattern_color = color_grid,
                    alpha = 0) +
  theme_bw() +
  guides(pattern = "none") + 
  ggtitle(paste0('Average posterior density for ',ns[1])) +
  xlab(ns[-1][[1]]) +
  ylab(ns[-1][[2]]) +
  labs(fill = "Posterior Mean")

plotsP <- ggplot(t1d, aes(mod1Val, mod2Val)) +
  geom_tile(aes(fill = bayes_p)) +
  scale_color_identity() +
  scale_fill_gradient2(low = color_low, 
                       high = color_high,
                       mid = color_mid,
                       midpoint = 0.5) +
  geom_tile_pattern(data = pat, aes(pattern = pattern),
                    pattern_density = grid_density,
                    pattern_spacing = grid_spacing,
                    pattern_color = color_grid,
                    alpha = 0) +
  theme_bw() +
  guides(pattern = "none") + 
  ggtitle(paste0('Bayesian p-value for ', ns[1])) +
  xlab(ns[-1][[1]]) +
  ylab(ns[-1][[2]]) +
  labs(fill = expression("Bayesian" ~ italic('p') * "-value"))

plotsMean
```


## Graphical Output and Customisation {#sec-graphics}

All plots are standard **ggplot2** objects [@Wickham2016] and can be further
modified using **ggplot2**'s layer system. The crosshatch pattern for
non-significant regions is produced by **ggpattern** [@ggpattern2022].
@tbl-colors documents the colour and pattern parameters shared across all
plotting functions. The colours for high, mid, and low point for continuous
plots are taken from the `Zissou1` palette from the **wesanderson** package
[@wesanderson]. Direct use of palettes is not supported but users can always
overlay their own color palettes post-hoc by calling
`scale_fill_gradient2()`/`scale_color_gradient2()` (for more and alternatives
see `?ggplot2::scale_fill_gradient()`).

```{r}
#| label: tbl-colors
#| tbl-cap: "Colour and pattern parameters available in all **int3ract**
#|  plotting functions. Default colour values are highlighted in their
#|   respective colour."
#| echo: false
#| cache: true

library(gt)

# Hex background and contrasting foreground for each colour-valued row.
# Rows without a colour default (grid_density, grid_spacing,
# crosshatch_non_sig) get NA and are left unstyled.
bg_cols <- c("#43CD80", "#D2691E", "#000000",
             "#3B9AB2", "#EBCC2A", "#F21A00",
             "#000000",  NA,        NA,
             "#666666",  NA)

fg_cols <- c("#000000", "#FFFFFF", "#FFFFFF",
             "#000000", "#000000", "#FFFFFF",
             "#FFFFFF",  NA,        NA,
             "#FFFFFF",  NA)

params <- data.frame(
  Parameter = c("sig_color", "non_sig_color", "line_color",
                "color_low", "color_mid",     "color_high",
                "color_grid", "grid_density", "grid_spacing",
                "color_values", "crosshatch_non_sig"),
  Description = c(
    "Ribbon fill for significant regions (2-way)",
    "Ribbon fill for non-significant regions (2-way)",
    "Point-estimate line colour (2-way)",
    "Low end of diverging fill scale (heatmaps)",
    "Midpoint colour of diverging fill scale",
    "High end of diverging fill scale",
    "Crosshatch line colour",
    "Density of crosshatch lines",
    "Spacing between crosshatch lines",
    "Within-cell text colour (heatmaps, \u2264 7 values)",
    "If TRUE, crosshatch marks the non-significant region"
  ),
  Default = c("'seagreen3'", "'chocolate'", "'black'",
              "'#F05039'",  "'#EBCC2A'",     "'#000066'",
              "'black'",    "0.01",         "0.1",
              "'grey40'",   "TRUE"),
  stringsAsFactors = FALSE
)

# Build the base gt table with monospace styling on Parameter and Default
tbl <- gt(params) |>
  tab_style(
    style    = cell_text(font = google_font("Source Code Pro"), size = "small"),
    locations = cells_body(columns = c(Parameter, Default))
  ) |>
  tab_style(
    style    = cell_text(font = google_font("Source Code Pro"), size = "small"),
    locations = cells_column_labels(columns = c(Parameter, Default))
  ) |>
  cols_width(
    Parameter   ~ px(160),
    Description ~ px(330),
    Default     ~ px(110)
  ) |>
  opt_row_striping()

# Apply per-row background + foreground to the Default cell
for (i in seq_len(nrow(params))) {
  if (!is.na(bg_cols[i])) {
    tbl <- tbl |>
      tab_style(
        style = list(
          cell_fill(color = bg_cols[i]),
          cell_text(color = fg_cols[i])
        ),
        locations = cells_body(columns = Default, rows = i)
      )
  }
}

# Add a thin border around the 'white' cell so it is visible
tbl <- tbl |>
  tab_style(
    style     = cell_borders(sides = "all", color = "#CCCCCC", weight = px(1)),
    locations = cells_body(columns = Default, rows = 5)   # row 5 = color_mid
  )

tbl
```
## Summary and Outlook {#sec-conc}

**int3ract** provides a unified interface for the Johnson--Neyman technique
and its three-way extension across a broad class of frequentist and Bayesian
models. By exposing the core computation in a single model-agnostic function,
`JNK_int()`, the package invites adoption beyond the two model classes for
which dedicated wrappers currently exist. Future development plans include:

1. A wrapper for `merMod` objects from **lme4** to allow for randomly varying
interactions.
2. A consideration for FDR-correction.
3. A general Bayesian application

Researchers are encouraged to report not only the JN boundaries but also the
full plots as supplementary material, in keeping with calls for more
transparent and complete reporting of interaction effects [@Simonsohn2020].

## Dependencies
To give due credit, next to the already listed packages, **int3ract** also
depends on **scales**, **tidyr**, and **tibble** [@scales; @tidy; @tibble].

## References {.unnumbered}

::: {#refs}
:::
